My title

### 2022-03-17 outputs/paper/paper.Rmd
added 3 graphs to results

# Load libraries
library(tidyverse)
library(kableExtra)

# Read in data
gss_data <- read_csv(here::here("inputs/data/gss.csv"))

@@

>>>>

# Load libraries
library(tidyverse)
library(kableExtra)
library(patchwork)

# Read in data
raw_data <- read_csv(here::here("inputs/data/AAae9eTh.csv"))
gss_data <- read_csv(here::here("inputs/data/gss.csv"))

@@

```{r householdwrangling, include=FALSE, message=FALSE, warning=FALSE}

# Data wrangling for household
household_data <- raw_data %>%
  select(hsdsizec, famtype) %>%
  mutate(household_size = case_when(hsdsizec == 6 ~ "6+",
                                    TRUE ~ as.character(hsdsizec)), 
         family_type = case_when(famtype == 1 ~ "Couple",
                                 famtype == 2 ~ "Intact family*",
                                 famtype == 3 ~ "Step-family with common child",
                                 famtype == 4 ~ "Step-family without a common child",
                                 famtype == 5 ~ "Single-parent family",
                                 famtype == 6 ~ "No spouse/partner or children in the
household"))

# get percents of each household size
household_percents <- household_data %>%
  group_by(household_size) %>%
  summarise(percent = 100*n()/nrow(household_data))

# 5+ people in a houshold percentage
five_plus_households <- household_percents %>%
  filter(household_size >= 5) %>%
  select(percent) %>%
  summarise(total_percent = sum(percent))

Figure \@ref(fig:householdsize) shows the distribution of household sizes with family type. The most common household sizes are 1 and 2-person, so Canadians generally live alone or in a romantic relationship. For household sizes greater than 2, they are mostly nuclear families, with 3-4 persons. Only `r round(five_plus_households$total_percent, 2)`% of households have 5 or more people living together.

```{r householdsize, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of family type by household size"}

# Plot for household size
household_data %>%
  ggplot(aes(x = household_size, fill = family_type)) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(breaks = seq(0, 8000, by = 1000)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(x = "Household size",
       y = "Number of responses",
       fill = "Family Type",
       caption = "*Intact family = family in which both biological\nparents are present in the home.")

## Marriage and Children

```{r numchildrenwrangling, include=FALSE, message=FALSE, warning=FALSE}

# Data wrangling for total number of children
child_data <- gss_data %>%
  mutate(total_children1 = case_when(total_children == 7 ~ "7+",
                                     TRUE ~ as.character(total_children)
                                     ))
# get percents of each number
child_percents <- child_data %>%
  group_by(total_children1) %>%
  summarise(percent = 100*n()/nrow(child_data))

# 5+ people in a houshold percentage
zero_children <- child_percents %>%
  filter(total_children1 == 0) %>%
  select(percent) %>%
  summarise(total_percent = sum(percent))

Figure \@ref(fig:numchildren) shows the distribution of the total number of children per family. Note that this is not *per household* i.e. persons living together, but the number of children ever had by the respondent of the household. The two tallest bars of nearly equal height are for 0 and 2 children, followed by 1 and 3 children also being similar in frequency, but about half as much as for 0 and 2. Thus, `r round(zero_children$total_percent,2)`% Canadians do not have any children and those who do usually have 1-3 children.

```{r numchildren, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of total number of children per family"}
# Plot for total number of children
child_data %>%
  select(total_children1) %>%
  ggplot(aes(x = total_children1)) +
  geom_bar(fill = "#377EB8") +
  scale_y_continuous(breaks = seq(0, 7000, by = 1000)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(title = "Distribution of total number of children per family",
       x = "Number of children",
       y = "Number of responses")
```

Figure \@ref(fig:agemarbirth) shows the distributions of the age of respondents when they first got married and had children.

```{r agemarbirth, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Distribution of ages at marriage and birth"}
# Plot for age at first marriage
marriage_plot<- gss_data %>%
  # mutate(age_at_first_marriage = round(age_at_first_marriage)) %>%
  select(age_at_first_marriage) %>%
  ggplot(aes(x = age_at_first_marriage)) +
  geom_histogram(bins = 65, fill = "#E41A1C") +
  scale_x_continuous(limits = c(15, 50),breaks = seq(0, 50, by = 10)) +
  scale_y_continuous(limits = c(0, 600),breaks = seq(0, 800, by = 100)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(title = "A. Distribution of age at first marriage",
       x = "Age at first marriage",
       y = "Number of responses")

# Plot for age at first birth
first_birth_plot <- gss_data %>%
  # mutate(age_at_first_birth = round(age_at_first_birth)) %>%
  select(age_at_first_birth) %>%
  ggplot(aes(x = age_at_first_birth)) +
  geom_histogram(bins = 65, fill = "#4DAF4A") +
 scale_x_continuous(limits = c(15, 50), breaks = seq(0, 50, by = 10)) +
  scale_y_continuous(limits = c(0, 600),breaks = seq(0, 800, by = 100)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  labs(title = "B. Distribution of age at first birth",
       x = "Age at first birth",
       y = "Number of responses")

marriage_plot / first_birth_plot
